---
title: "Data Viz Assignment"
author: "Alana Graham"
date: "October 9th, 2026"
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
    df-print: paged
execute:
  echo: true
  warning: false
---

## Story

Worldwide box office revenues were identified through https://www.boxofficemojo.com/year/world/.

Adjustments for inflation were made by referencing cumulative inflation figures from https://www.officialdata.org/us/inflation/.

I then used the Wayback Machine (https://web.archive.org/) to identify the Rotten Tomatoes critic and all-audience scores within the first 90 days of each movie's initial theatrical release.

I then identified whether movies were sequels or based on existing, well-known intellectual property at the time of theatrical release.

## Version A: Scientific

```{r, setup}
library(tidyverse)
library(gt)
library(psych)
library(readxl)
library(dplyr)
library(janitor)
library(ggplot2)
library(scales)
library(patchwork)
library(gganimate)
library(gifski)
library(readr)
library(ggiraph)
library(ggExtra)

Data_Viz_Assignment_Dataset <- read_xlsx("Data Viz Assignment Dataset.xlsx")
Top_Movies <- Data_Viz_Assignment_Dataset |> 
  select(Year:Genre) |> 
  filter(Year != "NA") |> 
  mutate(Year = as.numeric(Year))
  
COVID_Removed_List <- Top_Movies |> 
  filter(Year != "2020", Year != "2021")

## Graph 1: Revenue (Y) by Year (X)
  ## Scientific --> Scatter plot
Graph_1_Scientific <- ggplot(data = Top_Movies, aes(x = Year, y = `Inflation-Adjusted Box Office`))+
  geom_jitter(alpha = .4)+
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  theme_minimal()+
  labs(title = str_wrap("Inflation-Adjusted Box Office Revenue for the 5 Highest Grossing Films of 2006-2025"), x = "Year", y = "Inflation-Adjusted Worldwide Box Office (USD)")+
  scale_y_continuous(labels = label_dollar(scale_cut = cut_short_scale()))+
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5))

year_by_revenue_correlation <- cor.test(Top_Movies$Year, Top_Movies$`Inflation-Adjusted Box Office`, method = "pearson")

COVID_removed_year_by_revenue_correlation <- cor.test(COVID_Removed_List$Year, COVID_Removed_List$`Inflation-Adjusted Box Office`, method = "pearson")

Graph_1_Accessible <- ggplot(data = Top_Movies, aes(x = Year, y = `Inflation-Adjusted Box Office`))+
  stat_summary(aes(group = 1),
    fun = mean,
    geom = "line",
    colour = "red3",
    linewidth = 1.2) +
  stat_summary(aes(group = 1),
    fun = mean,
    geom = "point",
    colour = "orangered4",
    fill = "orangered3",
    shape = 21,
    size = 3) +
  annotate("segment",
    x = 2016,
    y = 750000000,
    xend = 2019.7,
    yend = 540000000,
    colour = "orangered3",
    linewidth = 0.8,
    arrow = arrow(length = unit(0.15, "inches"))) +
  annotate("label",
    x = 2016,
    y = 750000000,
    label = "COVID-19 pandemic",
    fill = "lemonchiffon2",
    colour = "black",
    label.size = 0,
    size = 4) +
  labs(title = "How Much Do the Biggest Movies Really Make?",
    subtitle = "Average worldwide box office for the five biggest films of 2006-2025, adjusted for inflation",
    x = "Year",
    y = "Average worldwide box office (USD)") +
  scale_y_continuous(labels = label_dollar(scale_cut = cut_short_scale())) +
  scale_x_continuous(breaks = seq(2006, 2025, 2)) +
  theme_classic() +
  theme(plot.background = element_rect(fill = "lemonchiffon3"),
    panel.background = element_rect(fill = "lemonchiffon3"),
    text = element_text(colour = "black"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.title = element_text(colour = "black"),
    plot.title = element_text(size = 18, face = "bold"))

  animated_plot <- Graph_1_Accessible + 
  transition_reveal(Year)
  
  ## Graph 2: Revenue (Y) by Critic and Audience Ratings (X) --> Although both critic and audience ratings are related to box office revenue, these ratings aren't very good at explaining box office turnout. Instead, we need to look at the factors that put butts into seats in the first place.
  ## Scientific --> Scatter plot disaggregated by rating type
Ratings_Revenue <- Top_Movies |>
pivot_longer(cols = c(`90-Day Rotten Tomatoes audience score`, `90-Day Rotten Tomatoes critic score`),
  names_to = "Rating Type",
  values_to = "Rating") |> 
  mutate(Rating = Rating*100)

Graph_2_Scientific <- ggplot(data = Ratings_Revenue, aes(x = Rating, y = `Inflation-Adjusted Box Office`, colour = `Rating Type`)) +
  geom_point(na.rm = TRUE) +
  geom_smooth(method = "lm",
  se = FALSE,
  fullrange = TRUE,
  linewidth = 1.2,
  na.rm = TRUE) +
labs(title = "Associations Between Critic/Audience Ratings and Box Office Revenue",
  x = "Rating (%)",
  y = "Box Office Revenue (USD)") +
scale_x_continuous(limits = c(18, 100),
  breaks = seq(20, 100, 20)) +
scale_y_continuous(labels = label_dollar(
  scale_cut = cut_short_scale())) +
scale_colour_grey(start = 0.2, end = 0.7) +
theme_minimal() +
  theme(legend.position = "bottom") +
  theme(text = element_text(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", linewidth = 0.5))+
annotate("label",
  x = 20,
  y = 2500000000,
  label = "Critic r = .21
Audience r = .12 ",
  fill = "white",
  colour = "black",
  label.size = 0,
  size = 4,
  hjust = 0) 
  
rating_by_revenue_correlation <- cor.test(Top_Movies$`90-Day Rotten Tomatoes critic score`, Top_Movies$`Inflation-Adjusted Box Office`, method = "pearson")

rating_by_audience_correlation <- cor.test(Top_Movies$`90-Day Rotten Tomatoes audience score`, Top_Movies$`Inflation-Adjusted Box Office`, method = "pearson")

critic_by_audience_correlation <- cor.test(Top_Movies$`90-Day Rotten Tomatoes audience score`, Top_Movies$`90-Day Rotten Tomatoes critic score`, method = "pearson")

## Accessible --> Patched scatter plots
  
Graph_2_Accessible_Audience <- ggplot(data = Top_Movies, aes(x = `90-Day Rotten Tomatoes audience score`, y = `Inflation-Adjusted Box Office`)) +
  geom_point_interactive(aes(tooltip = `Movie Title`, data_id = `Movie Title`), na.rm = TRUE, colour = "orangered3", size = 3, position = "jitter") +
  scale_x_continuous(limits = c(18, 100),
      breaks = seq(20, 100, 20)) +
  scale_y_continuous(labels = label_dollar(
    scale_cut = cut_short_scale())) +
  labs(x = "Audience Rating (%)",
       y = "Box Office Revenue (USD)") +
  theme_classic() +
  theme(text = element_text(colour = "black"),
        plot.background = element_rect(fill = "lemonchiffon3"),
        panel.background = element_rect(fill = "lemonchiffon3"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = "black"),
        axis.line = element_line(color = "black", linewidth = 0.5))

Graph_2_Accessible_Audience <- Graph_2_Accessible_Audience +
  geom_smooth(method = "lm",
                          se = FALSE,
                          fullrange = TRUE,
                          linewidth = 1.2,
                          na.rm = TRUE,
                          colour = "orangered3")

Graph_2_Accessible_Critic <- ggplot(data = Top_Movies, aes(x = `90-Day Rotten Tomatoes critic score`, y = `Inflation-Adjusted Box Office`)) +
  geom_point_interactive(aes(tooltip = `Movie Title`, data_id = `Movie Title`), na.rm = TRUE, colour = "orangered3", size = 3, position = "jitter") +
  scale_x_continuous(limits = c(18, 100),
       breaks = seq(20, 100, 20)) +
  scale_y_continuous(labels = label_dollar(
       scale_cut = cut_short_scale())) +
  labs(x = "Critic Rating (%)",
       y = "Box Office Revenue (USD)") +
  theme_classic() +
  theme(text = element_text(colour = "black"),
       plot.background = element_rect(fill = "lemonchiffon3"),
       panel.background = element_rect(fill = "lemonchiffon3"),
       panel.grid.major.x = element_blank(),
       panel.grid.minor = element_blank(),
       panel.grid.major.y = element_line(colour = "black"),
       axis.line = element_line(color = "black", linewidth = 0.5)) 

Graph_2_Accessible_Critic <- Graph_2_Accessible_Critic +
  geom_smooth(method = "lm",
                          se = FALSE,
                          fullrange = TRUE,
                          linewidth = 1.2,
                          na.rm = TRUE,
                          colour = "orangered3")

```

```{r, Graph_1_Scientific}
Graph_1_Scientific
```

```{r, Graph_2_Scientific}
Graph_2_Scientific
```

You can add options to executable code like this

## Version B: Accessible/Dynamic

```{r, accessible1}
animate(animated_plot, renderer = gifski_renderer())

```

Interactive Audience Plot

```{r, girafe}
girafe(
  ggobj = Graph_2_Accessible_Audience,
  options = list(
    opts_hover(css = "fill:black;stroke:black;cursor:pointer;"),
    opts_tooltip(css = "background-color:white;color:black;padding:5px;border-radius:3px;")
  )
)

```

Interactive Critic Plot

```{r}
girafe(
  ggobj = Graph_2_Accessible_Critic,
  options = list(
    opts_hover(css = "fill:black;stroke:black;cursor:pointer;"),
    opts_tooltip(css = "background-color:white;color:black;padding:5px;border-radius:3px;")
  )
)
```
