---
title: "Data Viz Assignment"
author: "Alana Graham"
date: "October 9 2026"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
    code-summary: "Show code"
    df-print: paged
execute:
  echo: true
  warning: false
---

## Story

(<https://www.reuters.com/business/media-telecom/us-box-office-is-booming-fewer-people-are-going-movies-2026-08-05/>)

Worldwide box office revenues were identified through <https://www.boxofficemojo.com/year/world/>.

Adjustments for inflation were made by referencing cumulative inflation figures from <https://www.officialdata.org/us/inflation/>.

I then used the Wayback Machine (<https://web.archive.org/>) to identify the Rotten Tomatoes critic and all-audience scores within the first 90 days of each movie's initial theatrical release.

I then identified whether movies were sequels or based on existing, well-known intellectual property at the time of theatrical release.

## Version A: Scientific

```{r, setup}
library(tidyverse)
library(gt)
library(psych)
library(readxl)
library(dplyr)
library(janitor)
library(ggplot2)
library(scales)
library(patchwork)
library(gganimate)
library(gifski)
library(readr)
library(ggiraph)
library(ggExtra)
library(plotly)

Data_Viz_Assignment_Dataset <- read_xlsx("Data Viz Assignment Dataset.xlsx")
Top_Movies <- Data_Viz_Assignment_Dataset |> 
  select(Year:Genre) |> 
  filter(Year != "NA") |> 
  mutate(Year = as.numeric(Year)) |> 
   mutate(`90-Day Rotten Tomatoes critic score` = `90-Day Rotten Tomatoes critic score`*100) |> 
  mutate(`90-Day Rotten Tomatoes audience score` = `90-Day Rotten Tomatoes audience score`*100)
  
COVID_Removed_List <- Top_Movies |> 
  filter(Year != "2020", Year != "2021")

## Graph 1: Revenue and Ratings (Y) by Year (X)
  ## Scientific --> Scatter plot
Graph_1_Scientific <- ggplot(data = Top_Movies, aes(x = Year, y = `Inflation-Adjusted Box Office`))+
  geom_jitter(alpha = .4)+
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  theme_minimal()+
  labs(title = str_wrap("Inflation-Adjusted Box Office Revenue for the 5 Highest Grossing Films of 2006-2025"), x = "Year", y = "Inflation-Adjusted Worldwide Box Office (USD)")+
  scale_y_continuous(labels = label_dollar(scale_cut = cut_short_scale()))+
  scale_x_continuous(limits = c(2006, 2025),
                     breaks = seq(2006, 2025, 5)) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5))

year_by_revenue_correlation <- cor.test(Top_Movies$Year, Top_Movies$`Inflation-Adjusted Box Office`, method = "pearson")

COVID_removed_year_by_revenue_correlation <- cor.test(COVID_Removed_List$Year, COVID_Removed_List$`Inflation-Adjusted Box Office`, method = "pearson")

Year_Ratings <- Top_Movies |>
  pivot_longer(cols = c(`90-Day Rotten Tomatoes audience score`, `90-Day Rotten Tomatoes critic score`),
               names_to = "Rating Type",
               values_to = "Rating")

Graph_1_Scientific_Rating <- ggplot(data = Year_Ratings, aes(x = Year, y = Rating, colour = `Rating Type`)) +
  geom_point(na.rm = TRUE) +
  geom_smooth(method = "lm",
              se = FALSE,
              fullrange = TRUE,
              linewidth = 1.2,
              na.rm = TRUE) +
  labs(title = "Rotten Tomatoes Ratings Have Steadily Increased Year-by-Year",
       x = "Year",
       y = "Rotten Tomatoes Rating (%)") +
  scale_y_continuous(limits = c(18, 100),
                     breaks = seq(20, 100, 20)) +
  scale_x_continuous(limits = c(2006, 2025),
                     breaks = seq(2006, 2025, 5)) +
  scale_colour_grey(start = 0.2, end = 0.7) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(text = element_text(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", linewidth = 0.5))+
  annotate("label",
           x = 2015,
           y = 30,
           label = "Critic r = .26
Audience r = .48 ",
           fill = "white",
           colour = "black",
           label.size = 0,
           size = 4,
           hjust = 0) 


## Accessible
Graph_1_Accessible <- ggplot(data = Top_Movies, aes(x = Year, y = `Inflation-Adjusted Box Office`))+
  stat_summary(aes(group = 1),
    fun = mean,
    geom = "line",
    colour = "red3",
    linewidth = 1.2) +
  stat_summary(aes(group = 1),
    fun = mean,
    geom = "point",
    colour = "orangered4",
    fill = "orangered3",
    shape = 21,
    size = 3) +
  annotate("segment",
    x = 2016,
    y = 750000000,
    xend = 2019.7,
    yend = 540000000,
    colour = "orangered3",
    linewidth = 0.8,
    arrow = arrow(length = unit(0.15, "inches"))) +
  annotate("label",
    x = 2016,
    y = 750000000,
    label = "COVID-19 pandemic",
    fill = "lemonchiffon2",
    colour = "black",
    label.size = 0,
    size = 4) +
  labs(title = str_wrap("How Much Do the Biggest Movies Really Make?"),
    subtitle = "Average worldwide box office for the five biggest films of 2006-2025, adjusted for inflation",
    x = "Year",
    y = "Average worldwide box office (USD)") +
  scale_y_continuous(labels = label_dollar(scale_cut = cut_short_scale())) +
  scale_x_continuous(breaks = seq(2006, 2025, 2)) +
  theme_classic() +
  theme(plot.background = element_rect(fill = "lemonchiffon3"),
    panel.background = element_rect(fill = "lemonchiffon3"),
    text = element_text(colour = "black"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.title = element_text(colour = "black"),
    plot.title = element_text(size = 18, face = "bold"))

  animated_plot_Graph_1_Box_Office <- Graph_1_Accessible + 
  transition_reveal(Year)

Graph_1_Accessible_Rating <- ggplot(data = Year_Ratings, aes(x = Year, y = Rating, colour = `Rating Type`, group = `Rating Type`)) +
  stat_summary(fun = mean,
               geom = "line",
               linewidth = 1.2) +
  stat_summary(fun = mean,
               geom = "point",
               shape = 21,
               size = 3) +
  labs(title = "How Much Do the Biggest Movies Really Make?",
       subtitle = "Average Rotten Tomatoes Ratings for the five biggest films of 2006-2025, adjusted for inflation",
       x = "Year",
       y = "Rotten Tomatoes Ratings (%))") +
  scale_y_continuous(limits = c(40, 100),
                     breaks = seq(40, 100, 20)) +
  scale_x_continuous(breaks = seq(2006, 2025, 2)) +
  scale_colour_manual(values = c(
      "90-Day Rotten Tomatoes audience score" = "orangered3",
      "90-Day Rotten Tomatoes critic score" = "steelblue4"),
    labels = c("90-Day Rotten Tomatoes audience score" = "Audience",
      "90-Day Rotten Tomatoes critic score" = "Critics")) +
  theme_classic() +
  theme(legend.background = element_rect(fill = "lemonchiffon3"), 
        plot.background = element_rect(fill = "lemonchiffon3"),
        panel.background = element_rect(fill = "lemonchiffon3"),
        text = element_text(colour = "black"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "black"),
        axis.text = element_text(colour = "black"),
        axis.title = element_text(colour = "black"),
        plot.title = element_text(size = 18, face = "bold"))

animated_plot_Graph_1_Ratings <- Graph_1_Accessible_Rating + 
  transition_reveal(Year)
  
  ## Graph 2: Revenue (Y) by Critic and Audience Ratings (X)
  ## Scientific --> Scatter plot disaggregated by rating type
Ratings_Revenue <- Top_Movies |>
pivot_longer(cols = c(`90-Day Rotten Tomatoes audience score`, `90-Day Rotten Tomatoes critic score`),
  names_to = "Rating Type",
  values_to = "Rating")

Graph_2_Scientific <- ggplot(data = Ratings_Revenue, aes(x = Rating, y = `Inflation-Adjusted Box Office`, colour = `Rating Type`)) +
  geom_point(na.rm = TRUE) +
  geom_smooth(method = "lm",
  se = FALSE,
  fullrange = TRUE,
  linewidth = 1.2,
  na.rm = TRUE) +
labs(title = str_wrap("The Financial Success of BlockBusters Have Little To Do With Their Critic and Audience Ratings"),
  x = "Rating (%)",
  y = "Box Office Revenue (USD)") +
scale_x_continuous(limits = c(18, 100),
  breaks = seq(20, 100, 20)) +
scale_y_continuous(labels = label_dollar(
  scale_cut = cut_short_scale())) +
scale_colour_grey(start = 0.2, end = 0.7) +
theme_minimal() +
  theme(legend.position = "bottom") +
  theme(text = element_text(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", linewidth = 0.5))+
annotate("label",
  x = 20,
  y = 2500000000,
  label = "Critic r = .21
Audience r = .12 ",
  fill = "white",
  colour = "black",
  label.size = 0,
  size = 4,
  hjust = 0) 
  
rating_by_revenue_correlation <- cor.test(Top_Movies$`90-Day Rotten Tomatoes critic score`, Top_Movies$`Inflation-Adjusted Box Office`, method = "pearson")

rating_by_audience_correlation <- cor.test(Top_Movies$`90-Day Rotten Tomatoes audience score`, Top_Movies$`Inflation-Adjusted Box Office`, method = "pearson")

critic_by_audience_correlation <- cor.test(Top_Movies$`90-Day Rotten Tomatoes audience score`, Top_Movies$`90-Day Rotten Tomatoes critic score`, method = "pearson")

## Accessible --> Patched scatter plots
  
Graph_2_Accessible_Audience <- ggplot(data = Top_Movies, aes(x = `90-Day Rotten Tomatoes audience score`, y = `Inflation-Adjusted Box Office`)) +
  geom_point_interactive(aes(tooltip = `Movie Title`, data_id = `Movie Title`), na.rm = TRUE, fill = "orangered3", size = 4, position = "jitter", alpha = 0.8, shape = 21, colour = "black", stroke = 0.5) +
  scale_x_continuous(limits = c(18, 100),
      breaks = seq(20, 100, 20)) +
  scale_y_continuous(labels = label_dollar(
    scale_cut = cut_short_scale())) +
  labs(x = "Audience Rating (%)",
       y = "Box Office Revenue (USD)") +
  theme_classic() +
  theme(axis.text = element_text(size = 18),
    axis.title = element_text(size = 18),
        plot.background = element_rect(fill = "lemonchiffon3"),
        panel.background = element_rect(fill = "lemonchiffon3"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = "black"),
        axis.line = element_line(color = "black", linewidth = 0.5))

Graph_2_Accessible_Audience <- Graph_2_Accessible_Audience +
  geom_smooth(method = "lm",
                          se = FALSE,
                          fullrange = TRUE,
                          linewidth = 1.2,
                          na.rm = TRUE,
                          colour = "grey23")

Graph_2_Accessible_Critic <- ggplot(data = Top_Movies, aes(x = `90-Day Rotten Tomatoes critic score`, y = `Inflation-Adjusted Box Office`)) +
  geom_point_interactive(aes(tooltip = `Movie Title`, data_id = `Movie Title`), na.rm = TRUE, fill = "orangered3", size = 4, position = "jitter", alpha = 0.8, shape = 21, colour = "black", stroke = 0.5) +
  scale_x_continuous(limits = c(18, 100),
       breaks = seq(20, 100, 20)) +
  scale_y_continuous(labels = label_dollar(
       scale_cut = cut_short_scale())) +
  labs(x = "Critic Rating (%)",
       y = "Box Office Revenue (USD)") +
  theme_classic() +
  theme(axis.text = element_text(size = 18),
    axis.title = element_text(size = 18),
       plot.background = element_rect(fill = "lemonchiffon3"),
       panel.background = element_rect(fill = "lemonchiffon3"),
       panel.grid.major.x = element_blank(),
       panel.grid.minor = element_blank(),
       panel.grid.major.y = element_line(colour = "black"),
       axis.line = element_line(color = "black", linewidth = 0.5)) 

Graph_2_Accessible_Critic <- Graph_2_Accessible_Critic +
  geom_smooth(method = "lm",
                          se = FALSE,
                          fullrange = TRUE,
                          linewidth = 1.2,
                          na.rm = TRUE,
                          colour = "grey23")


## Rendering Interactive Plots

Combined_Graph_2 <- Graph_2_Accessible_Audience / Graph_2_Accessible_Critic + 
  plot_annotation(title = "Better Ratings Don't Guarantee Higher Earnings",
    subtitle = "Click on the dots to see the movie title!") +
  theme(plot.title = element_text(size = 30, hjust = 0, colour = "black"),
    plot.background = element_rect(fill = "lemonchiffon3", colour = NA))

### Graph 3: Ratings and Box Office (Y) by Genre (X)
## Scientific

Ratings_by_Genre <- Top_Movies |> 
  group_by(Genre) |> 
  summarise(Audience = mean(`90-Day Rotten Tomatoes audience score`, na.rm = TRUE), Critic = mean(`90-Day Rotten Tomatoes critic score`, na.rm = TRUE), 'Box Office' = mean(`Inflation-Adjusted Box Office`, na.rm = TRUE), count = n()) |> 
  filter(count > 5) 

Graph_3_Scientific_Ratings <- ggplot(Ratings_by_Genre) +
  geom_segment(aes(x = Audience, xend = Critic, y = Genre, yend = Genre), colour = "darkgrey") +
  geom_point(aes(x = Audience, y = Genre, colour = "Audience"), 
    size = 3) +
  geom_point(aes(x = Critic, y = Genre, colour = "Critic"), 
    size = 3) +
  scale_colour_manual(name = "Rating Type",
    values = c("Critic" = "gray70", "Audience" = "black")) +
  labs(x = "Rotten Tomatoes Ratings") +
  theme_minimal() +
  theme(legend.position = "bottom",
    text = element_text(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5),
    axis.text = element_text(size = 10))

Key_Genres <- Top_Movies |> 
  filter(Genre == "Action" | Genre == "Adventure" | Genre == "Fantasy" | Genre == "Kids & Family" | Genre == "Sci-Fi") 

Graph_3_Scientific_Box_Office <- ggplot(Key_Genres, aes(x = Genre, y = `Inflation-Adjusted Box Office`)) +
  geom_violin(trim = TRUE, fill = "gray95", color = "black", alpha = 0.4) +
  geom_jitter(alpha = 0.4) +
  stat_summary(fun = "mean", 
               geom = "crossbar", 
               width = 0.5,
               colour = "black",
               linewidth = 0.4) +
  scale_y_continuous(labels = label_dollar(scale_cut = cut_short_scale())) +
  theme_minimal() +
  labs(y = "Average worldwide box office (USD)") +
  theme(text = element_text(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", linewidth = 0.5))

## Accessible

Year_Ratings_Genre <- Year_Ratings |> 
  filter(Genre == "Action" | Genre == "Adventure" | Genre == "Fantasy" | Genre == "Kids & Family" | Genre == "Sci-Fi")

Graph_3_Accessible_Ratings <- ggplot(Year_Ratings_Genre, aes(x = Genre, y = Rating, fill = `Rating Type`)) +
  stat_summary(aes(group = `Rating Type`),
    fun = mean,
    geom = "col",
    position = position_dodge(width = 0.85),
    width = 0.7,
    na.rm = TRUE) +
  geom_point_interactive(aes(fill = `Rating Type`, tooltip = `Movie Title`, data_id = `Movie Title`),
    shape = 21,
    stroke = 0.5,
    colour = "black",
    position = position_dodge(width = 0.8),
    alpha = 0.6,
    size = 2) +
  scale_y_continuous(limits = c(0, 100),
    breaks = seq(0, 100, 20)) +
  scale_fill_manual(values = c("90-Day Rotten Tomatoes critic score" = "orangered4",
      "90-Day Rotten Tomatoes audience score" = "grey23")) +
  labs(x = "Genre",
    y = "Rotten Tomatoes Score (%)",
    fill = "Rating Type") +
  theme_classic() +
  theme(text = element_text(colour = "black"),
        plot.background = element_rect(fill = "lemonchiffon3"),
        panel.background = element_rect(fill = "lemonchiffon3"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = "black"),
        axis.line = element_line(color = "black", linewidth = 0.5),
        legend.background = element_rect(fill = "lemonchiffon3", colour = NA),
        legend.position = "bottom")

Graph_3_Accessible_Box_Office <- ggplot(Ratings_by_Genre, aes(x = Genre, y = `Box Office`)) +
  geom_col(color = "black", fill = "orangered4") +
  scale_y_continuous(labels = label_dollar(scale_cut = cut_short_scale())) +
  theme_minimal() +
  labs(y = "Average worldwide box office (USD)") +
  theme_classic() +
          theme(text = element_text(colour = "black"),
                plot.background = element_rect(fill = "lemonchiffon3"),
                panel.background = element_rect(fill = "lemonchiffon3"),
                panel.grid.major.x = element_blank(),
                panel.grid.minor = element_blank(),
                panel.grid.major.y = element_line(colour = "black"),
                axis.line = element_line(color = "black", linewidth = 0.5)) 

Combined_Graph_3 <- Graph_3_Accessible_Ratings / Graph_3_Accessible_Box_Office + 
  plot_annotation((title = "Adventure Movies Tank in Critic Ratings but Score at the Box Office"),
                  theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0, colour = "black"),
                                plot.background = element_rect(fill = "lemonchiffon3", colour = NA)))


```

```{r, Graph_1_Scientific}
Graph_1_Scientific

Graph_1_Scientific_Rating
```

Notably, audiences consistently offered higher ratings than critics, with a marked absence of low ratings since 2020. Although both critic and audience ratings are related to box office revenue, these ratings aren't very good at explaining box office turnout. Instead, we need to look at a factor that initially draws audiences to the theatre.

```{r, Graph_2_Scientific}
Graph_2_Scientific
```

Although Sci-Fi movies achieved the highest average ratings and box office revenue, films that primarily fell under the genre of "Adventure" were simultaneously less likely to receive higher ratings from both critics and audiences (with larger discrepancies), yet no less successful at the box office. Ultimately, it's clear that what defines a "successful" film has more to do with one's definition of success (i.e., ratings, revenue, etc.) than any single characteristic of a film.

Too few movies in the categories of:

```{r, Scientific_Genre}
Graph_3_Scientific_Ratings / Graph_3_Scientific_Box_Office +
  plot_annotation((title = str_wrap("Genre Differentially Influences Ratings and Box Office Revenue")),
  theme = theme(plot.title = element_text(size = 16, hjust = 0, colour = "black")))
```

Importantly, revenue figures do not consider the roles of budgets or profits, which may be better indicators of "success". They also do not reflect the reality that some of these films would be distributed to a narrower selection of theatres or countries. Additionally, despite the fact that films based on well-known intellectual property (e.g., sequels) did not differ in their box office revenues or Rotten Tomatoes critic and audience ratings, only 13 of the 100 films in the current sample did not fall under this category. This may suggest that (1) the film industry disproportionately relies on established intellectual property because of greater perceived confidence in its resulting success and/or (2) the sample of 13 non-legacy films presented here are an atypically successful subset compared to movies based on established IP.

To best tease apart the factors contributing to various reflections of success, comparisons between the highest-earning movies of each year should also be made with the larger pool of films playing in theatres.

## Version B: Accessible/Dynamic

“A Box Office Hit, a Critical Darling, or a Crowd-Pleaser?”

```{r, accessible1}

animate(animated_plot_Graph_1_Box_Office, renderer = gifski_renderer())

animate(animated_plot_Graph_1_Ratings, renderer = gifski_renderer())
```

Although both critic and audience ratings are related to box office revenue, these ratings aren't very good at explaining box office turnout. Instead, we need to look at a factor that puts butts into seats in the first place: Genre

Interactive Audience Plot

```{r, interactive_audience_plot}
girafe(ggobj = Combined_Graph_2,
  width_svg = 12, height_svg = 12,
  options = list(opts_hover(css = "fill:black;stroke:black;cursor:pointer;"),
    opts_tooltip(css = "background-color:white;color:black;padding:5px;border-radius:3px;")))

```

Although Sci-Fi movies achieved the highest average ratings and box office revenue, films that primarily fell under the genre of "Adventure" were simultaneously less likely to receive higher ratings from both critics and audiences (with larger discrepancies), yet no less successful at the box office. Ultimately, it's clear that what defines a "successful" film has more to do with one's definition of success (i.e., ratings, revenue, etc.) than any single characteristic of a film.

Too few movies in the categories of:
```{r, Focus on Genre}
girafe(ggobj = Combined_Graph_3,
       width_svg = 10, height_svg = 12,
       options = list(opts_hover(css = "fill:black;stroke:black;cursor:pointer;"),
                      opts_tooltip(css = "background-color:white;color:black;padding:5px;border-radius:3px;")))
```

Importantly, revenue figures do not consider the roles of budgets or profits, which may be better indicators of "success". They also do not reflect the reality that some of these films would be distributed to a narrower selection of theatres or countries. Additionally, despite the fact that films based on well-known intellectual property (e.g., sequels) did not differ in their box office revenues or Rotten Tomatoes critic and audience ratings (not depicted here), only 13 of the 100 films in the current sample did not fall under this category. This may suggest that (1) the film industry disproportionately relies on established intellectual property because of greater perceived confidence in its resulting success and/or (2) the sample of 13 non-legacy films presented here are an atypically successful subset compared to movies based on established IP.

To best tease apart the factors contributing to various reflections of success, appropriate statistical comparisons between the highest-earning movies of each year should also be made with the larger pool of films playing in theatres.
